Reinforcement Learning Produces Dominant Strategies for the Iterated Prisoner's Dilemma

### 2017-07-08 main.tex
Up size of one of the FSM diagrams slightly.
Add brief description of previous tournaments.
Please modify etc.

% TODO: discuss earlier tournaments

In this work we describe how collections of strategies in the Axelrod library
have been used to train new strategies specifically to win IPD tournaments.
These strategies are trained using generic strategy archetypes based on e.g.
finite state
machines, arriving at particularly effective parameter choices through
evolutionary or particle swarm algorithms. There are several
previous publications that use evolutionary algorithms to
evolve IPD strategies in various circumstances
\cite{ashlock2006training, ashlock2015multiple, ashlock2006changes,
      ashlock2014shaped, ashlock2014evolution, barlow2015varying,
      fogel1993evolving, marks1989niche, sudo2015effects,
      vassiliades2010multiagent}. See also \cite{gaudesi2016exploiting} for a
strategy trained to win against a collection of well-known IPD opponents and see
\cite{franken2005particle} for a prior use of particle swarm algorithms. Our
results are unique in that we are able to train against a large collection of
well-known strategies available in the scientific literature. Crucially, the
software used in this work is openly available and can be used to train strategies
in the future in a reliable manner, with confidence that the opponent strategies
are correctly implemented, tested and documented.
Moreover, as of the time of writing,
we claim that this work contains the best known strategies for the iterated
prisoner's dilemma.

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.5\textwidth}
        \includestandalone[height=.3\textheight]{./assets/FSM16}
        \caption{Evolved\_FSM\_16: trained to maximise score in a standard
        tournament}
        \label{fig:fsm16}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includestandalone[height=.3\textheight]{./assets/FSM16noise}
        \caption{Evolved\_FSM\_16\_Noise\_05: trained to maximise score in a
        noisy tournament}
        \label{fig:fsm16noise}
    \end{subfigure}%
    \caption{Trained sixteen state finite state machine players}
    \label{fig:fsm_images}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_20_EvolvedLookerUp2_2_2_array.pdf}
        \caption{Cooperation rate per round for EvolvedLookerUp\_2\_2\_2 (over
        20 repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_20_Evolved_HMM_5_array.pdf}
        \caption{Cooperation rate per round for Evolved\_HMM\_5 (over 20
        repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_20_Evolved_FSM_16_array.pdf}
        \caption{Cooperation rate per round for Evolved\_FSM\_16 (over 20
        repetitions)}
    \end{subfigure}%

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_20_DBS_0-75_3_4_3_5_array.pdf}
        \caption{Cooperation rate per round for DBS (over 20 repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_20_Evolved_ANN_5_Noise_05_array.pdf}
        \caption{Cooperation rate per round for Evolved\_ANN\_5\_Noise\_05 (over 20
        repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_20_Evolved_FSM_16_Noise_05_array.pdf}
        \caption{Cooperation rate per round for Evolved\_FSM\_16\_Noise\_05 (over 20
        repetitions)}
    \end{subfigure}%

>>>>

Since Robert Axelrod's seminal tournament \cite{Axelrod1980a}, a number of
IPD\@ tournaments have been undertaken and are summarised in
Table~\ref{tbl:tournaments}. Further to the work described in
\cite{knight2016open} a regular set of standard, noisy and probabilistic ending
tournaments are carried out as more strategies are added to the Axelrod library.
Details and results are available here:
\url{http://axelrod-tournament.readthedocs.io}. This work presents a detailed
analysis of a tournament with 176 strategies, more details about this are given
in Section~\ref{sec:results}.

\begin{table}[!hbtp]
    \begin{center}
        \begin{tabular}{ccccc}
            \toprule
            Year     & Reference                  & Number of Strategies & Type     & Source Code\\
            \midrule
            1979     & \cite{Axelrod1980a}        & 13                   & Standard & Not immediately available\\
            1979     & \cite{Axelrod1980b}        & 64                   & Standard & Available in FORTRAN\\
            1991     & \cite{Bendor1991}          & 13                   & Noisy    & Not immediately available\\
            2002     & \cite{Stephens2002}        & 16                   & Wildlife & Not applicable\\
            2005     & \cite{kendall2007iterated} & 223                  & Varied   & Not available \\
            2012     & \cite{Stewart2012}         & 13                   & Standard & Not fully available \\
            2016     & \cite{knight2016open}       & 129                  & Standard & Fully available \\
            \bottomrule
        \end{tabular}
    \end{center}
    \caption{An overview of a selection of published tournaments. Not all
             tournaments were `standard' round robins; for more details
             see the indicated references.}\label{tbl:tournaments}
\end{table}

In this work we describe how collections of strategies in the Axelrod library
have been used to train new strategies specifically to win IPD tournaments.
These strategies are trained using generic strategy archetypes based on e.g.\
finite state
machines, arriving at particularly effective parameter choices through
evolutionary or particle swarm algorithms. There are several
previous publications that use evolutionary algorithms to
evolve IPD strategies in various circumstances
\cite{ashlock2006training, ashlock2015multiple, ashlock2006changes,
      ashlock2014shaped, ashlock2014evolution, barlow2015varying,
      fogel1993evolving, marks1989niche, sudo2015effects,
      vassiliades2010multiagent}. See also \cite{gaudesi2016exploiting} for a
strategy trained to win against a collection of well-known IPD opponents and see
\cite{franken2005particle} for a prior use of particle swarm algorithms. Our
results are unique in that we are able to train against a large collection of
well-known strategies available in the scientific literature. Crucially, the
software used in this work is openly available and can be used to train strategies
in the future in a reliable manner, with confidence that the opponent strategies
are correctly implemented, tested and documented.
Moreover, as of the time of writing,
we claim that this work contains the best known strategies for the iterated
prisoner's dilemma.

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.5\textwidth}
        \includestandalone[height=.3\textheight]{./assets/FSM16}
        \caption{Evolved\_FSM\_16: trained to maximise score in a standard
        tournament}
        \label{fig:fsm16}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.5\textwidth}
        \centering
        \includestandalone[height=.35\textheight]{./assets/FSM16noise}
        \caption{Evolved\_FSM\_16\_Noise\_05: trained to maximise score in a
        noisy tournament}
        \label{fig:fsm16noise}
    \end{subfigure}%
    \caption{Trained sixteen state finite state machine players}
    \label{fig:fsm_images}
\end{figure}

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_EvolvedLookerUp2_2_2_array.pdf}
        \caption{Cooperation rate per round for EvolvedLookerUp\_2\_2\_2 (over
        1000 repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_Evolved_HMM_5_array.pdf}
        \caption{Cooperation rate per round for Evolved\_HMM\_5 (over 1000
        repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_Evolved_FSM_16_array.pdf}
        \caption{Cooperation rate per round for Evolved\_FSM\_16 (over 1000
        repetitions)}
    \end{subfigure}%

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_DBS_0-75_3_4_3_5_array.pdf}
        \caption{Cooperation rate per round for DBS (over 1000 repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_Evolved_ANN_5_Noise_05_array.pdf}
        \caption{Cooperation rate per round for Evolved\_ANN\_5\_Noise\_05 (over 1000
        repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_Evolved_FSM_16_Noise_05_array.pdf}
        \caption{Cooperation rate per round for Evolved\_FSM\_16\_Noise\_05 (over 1000
        repetitions)}
    \end{subfigure}%
